---
title: "Demo Rmarkdown File"
output:
  pdf_document:
    keep_tex: true
header-includes:
- \usepackage{dcolumn}
- \usepackage{orcidlink}
- \usepackage{natbib}
- \bibliographystyle{abbrvnat}
- \setcitestyle{authoryear, open={((},close={)}}
---

```{r setup, warning = FALSE, message = FALSE}
library(tidyverse)
library(kableExtra)
library(palmerpenguins)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.height = 2, fig.width = 2)
```

# ORCID link

Kristen Hunter\orcidlink{0000-0002-5678-4620}

# Citations using natbib

This is a standard citation \citep{Raudenbush2002}.

\citet{Spybrook2014} is an in-text citation.

# Default table

```{r}
library(palmerpenguins)

knitr::kable(penguins[1:5, 1:5], caption = "Penguins!")
```


# Some table options

Note: it used to be `booktabs = TRUE` made prettier tables by default, but they seem to be default in the newest version of the package.

```{r}
# change NAs to print as dashed lines
options(knitr.kable.NA = '-')

# print table
knitr::kable(penguins[1:5, 1:5], digits = 3, booktabs = TRUE, 
             position = "h!", 
             caption = "Penguins!") %>%
  kableExtra::kable_styling(position = "left", font_size = 7)
```


# Citing tables

```{r penguinTable}
knitr::kable(penguins[1:5, 1:5], caption = "Penguins!") %>%
  kableExtra::kable_styling(position = "center")
```

Table \ref{tab:penguinTable} is a table of penguins.

# Citing figures

```{r penguinPlot, fig.cap="\\label{fig:penguinPlot}Penguin Bills", fig.pos = "h!"}
plot <- ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
print(plot)
```

Figure \ref{fig:penguinPlot} is a figure of penguins.

\clearpage
# Generating tables in a loop

```{r results = 'asis'}
for (i in 1:2)
{
  print(knitr::kable(penguins[i,1:3], caption = paste("Penguins", i)))
  cat("\\vspace{10mm}")
}
```

# Stargazer

```{r results = 'asis'}
library(stargazer)
# run regressions
linear.1 <- lm(
  rating ~ complaints + privileges + learning + raises + critical,
  data = attitude
)
linear.2 <- lm(
  rating ~ complaints + privileges + learning, data = attitude
)
stargazer(linear.1, linear.2, title = "Results", align = TRUE, header = FALSE)
```

\clearpage
\bibliography{refs.bib}

