---
title: "An Experimental Planning Analysis Exercise"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE
)

library( PUMP )
```

# Introduction

You are designing a randomized control trial for an educational program.
First, you would like to determine what your current study design will be able 
to detect.
If this is insufficient, you would like to know the necessary sample size to 
achieve a desired level of precision.
You also want to explore how different choices of parameter values will impact 
your findings.

Background information:

- Your trial will test $4$ different outcomes, which you think are moderately 
correlated (`rho = 0.4`).
- Your current plan is to do a cluster design, which means that entire 
districts will be assigned to either treatment or control 
(randomization occurs at level 3). This corresponds to plan code `d_m = 'd3.3_m3rc2rc'`.
- 50\% of the units will receive the treatment (`Tbar = 0.5`).
- You plan to use a Holm adjustment to adjust for the multiple outcomes (`HO`).
- You think there is going to be a treatment effect of about $0.25$ effect 
size units on all outcomes.

Currently, you have:

- 12 school districts
- 4 schools per district
- 500 students per school

You also have a few other pieces of information:

- You have 4 covariates describing the students, 2 covariates describing 
the schools, and no covariates at the district level. This corresponds to: `numCovar.1 = 4, numCovar.2 = 2, numCovar.3 = 0`. You think that the 
student-level covariates are quite predictive of the outcomes, but your school-level covariates are not very predictive of the outcomes. You pick
the explanatory values of the covariates at each level to be 
`R2.1 = 0.6, R2.2 = 0.2, R2.3 = 0`.
- You expect most of the variation to occur at the student level, not very 
much variation to occur at the school level, and some variation to be at the 
district level. In other words, schools within the same district should have 
similar outcomes, but outcomes may vary moderately between districts. Based
on these assumptions, you set `ICC.2 = 0.05, ICC.3 = 0.1`.

## Exercises

1. What is the minimal detectable effect size for your study if you want an 80% chance of detecting at least two of your four outcomes?


```{r}
library( PUMP )
# some code
```

2. You want to update your assumption concerning the predictive power of your 
covariates. What happens to MDES if the level 1 covariates are highly 
predictive of the first two outcomes, but not very predictive of the second two outcomes? We can set this using `R2.1 = c( 0.8, 0.8, 0.05, 0.05 )`.

```{r}
# some code

```

3.  Determine the number of school districts you would need with your 
original MDES of 0.25 to achieve 80% *individual* power.

```{r}
# some code

```

4. What happens to your needed sample size (number of school districts) if 
2 of your 4 outcomes are not actually impacted by treatment?

```{r}
# some code

```


5. Extend the above to check how number of schools districts needed 
would vary over a range of values for the intraclass correlation at 
levels 2 and 3 (ICC.2 and ICC.3).

```{r}
# some code

```

6. Visualize this change due to ICC.

```{r}
# some code

```

7. BONUS: What happens to the MDES if you change your design to randomize at the 
school level instead? (Note: you will have to pick a new model, choose what 
you want!)

```{r}
# some code

```
